---
title: "R Data types and Data structures"
author: "Nikhil Sawal"
date: "August 11, 2018"
output: rmarkdown::github_document
---

# 1. Data Types

### 1.1 Numberic [floating point values]
```{r}
a <- 2.2
```
### 1.2 Integer
```{r}
b <- 2
```

### 1.3 Logical [True/False]
```{r}
c <- TRUE
```

### 1.4 Characters [String]
##### Can use single or double quotes ["" / ""]
```{r}
d <- "Hello World!"
```  


# 2. Function class() - Checks Datatype
```{r}
class(a)
class(b)
class(c)
class(d)
```
**NOTE:** `b`, inspite of being an interger is stored as numeric, because all integers are stored as numeric

# 3. Vector
A vector is a one-dimensional array, that can hold character, logical or numeric data. 
We use the combine funtion to create a vector.

### 3.1 Numeric Vector
```{r}
num <- c(1,2,3,4,5)                # c() represents the combine function

class(num)
```
Since every element in the above vector is numeric the class of `num` is "numeric".  

### 3.2 Character and Boolean vector
```{r}
char <- c("x", "y", "z")
class(char)

bool <- c(T, F)
class(bool)
```  

Key Points about vectors  

* All elements within a vector should be of the **SAME DATA TYPE**
* If the elements belong to some other data type, R will convert their class. This 
  particular situation is referred to as coercion
```{r}
f <- c(TRUE, 1, 4, 5)
print(f)
```
Since the other elements in the vector f are numeric, TRUE is converted to numeric  

* True is converted to 1
* False is converted to 0  

Now,if we check the class of the vector `f`, it should be numeric.
```{r}
class(f)
```


Another example
```{r}
g <- c("afd", 1, 3, 6)
print(g)
class(g)
```
In case of the vector `g`, it is not possible to convert the character "afd" to a 
numeric type, so the other elements are converted to a character data type

### 3.3 Assign names to vector elements using **names()** function
```{r}
temp <- c(32,33,36,42,37,41,33)
names(temp) <- c("Mon", "Tue", "Wed", "Thurs", "Fri", "Sat", "Sun")
print(temp)
```
To make the above code reusable, assign the vector of days to a variable `days`. This would spare a lot of time for productive more work. 
```{r}
days <- c("Mon", "Tue", "Wed", "Thurs", "Fri", "Sat", "Sun")
names(temp) <- days
print(temp)
```
# 4. Vector Operations
### 4.1 Elementwise operations.
```{r}
v1 <- c(1,2,3)
v2 <- c(3,4,5)
```
### 4.2 Elementwise Addition
```{r}
print(v1 + v2)
```
### 4.3 Elementwise Subtraction
```{r}
print(v1 - v2)
```
### 4.4 Elementwise Division
```{r}
print(v1 / v2)
```
### 4.4 Elementwise Multiplication
```{r}
print(v1 * v2)
```

### 4.5 Vector built-in functions
```{r}
sum(v1)              # Returns Sum
mean(v2)             # Returns Mean
sd(v1)               # Returns Standard Deviation
max(v2)              # Returns Maximum
min(v1)              # Returns Minimum
prod(v2)             # Returns Product
length(v1)           # Returns Length
```

# 5. Indexing vector and slicing
```{r}
v1 <- c(1,2,3,4,6)
v2 <- c('a','b','c','d','e')
```

Indexing in R starts at `1`, unlike other programming languages, where it starts at `0`

### 5.1 Index a single element
```{r}
v1[1]
```

### 5.2 To grab the first and second element
```{r}
v1[c(1,2)]
v1[1:2]
```

### 5.3 To get elements in continuous sequence
```{r}
v1[3:5]
```

### 5.4 Indexing using names
```{r}
v <- c(1,2,3,4)
names(v) <- c('w', 'x', 'y', 'z')

v[2]
v['x']
v[c('x','y','z')]
```

### 5.5 Indexing or filtering by comparison or using logical statements

#### 5.5.1 Index all elements greater than 2
```{r}
print(v[v>2])
print(v>2)
```
Notice the difference b/w the above two statements

#### 5.5.2 Comparison operators - Evaluates True/False
```{r}
3 > 4
2 < 9
5 >= 5
2 <= 4
2 == 6
2 !=2
```
**Note:** that `'='` is an assignment operator and **CANNOT** be used for comparison
So, using `'='` for comparison will throw an error.

# 6. Matrix
A matrix is a 2-dimensional data structure which consists of elements of the same data type 

### 6.1 Create Matrix with `matrix()` function
```{r}
vec <- 1:10
matrix(vec)
```
This will create a matrix of 10 rows and 1 column, which is the default setting.

### 6.2 To specify the number of rows or columns, use the `nrow()` or `ncol()` argument respectively.

```{r}
matrix(vec, nrow = 2)
```

The default for filling the elements in a matrix is by columns. To change it to rows,
set the `byrow` argument to `TRUE`
```{r}
matrix(1:12, byrow = T, nrow = 4)
matrix(1:12, byrow = F, nrow = 4)
```
**Notice:** the difference. In the first matrix the elements are filled by rows since we set the `byrow` argument to `TRUE`. In the second matrix, the elements are filled by columns.

Let's do an example of creating a matrix, for the prices of apples and oranges for
weekdays. We start with creating 2 vectors `apples and oranges` and the elements of each vector, would be their prices for that respective weekdays.
```{r}
apples <- c(12, 11, 10, 9, 14)
oranges <- c(6, 9, 7, 11, 8)
```


Using the combine function `c()`, will create a vector
```{r}
prices <- c(apples, oranges)
prices
```

In order to create a matix, we perform the following operations.

```{r}
prices.matrix <- matrix(prices, byrow = T, nrow = 2)
print(prices.matrix)
```

Now, we can use the `colnames()` and `rownames()` function to give names to the columns and rows respectively.

```{r}
days <- c("Mon", "Tue", "Wed", "Thur", "Fri")
fruits <- c("Apples", "Oranges")

colnames(prices.matrix) <- days
rownames(prices.matrix) <- fruits

print(prices.matrix)
```

# 7. Matrix Arithmetic
We first create a matrix, `mat`.
```{r}
mat <- matrix(1:25, byrow = T, nrow = 5)
```

### 7.1 Matrix operations [ELEMENTWISE]
#### 7.1.1 Multiply all the elements by 2

```{r}
mat*2
```

#### 7.1.2 Divide all elements by 2

```{r}
mat/2
```

#### 7.1.3 Power - Square all elements in the matrix

```{r}
mat^2
```

#### 7.1.4 Using the comparison operator will return a matrix of logical values.
```{r}
mat > 14
``` 
#### 7.1.5 For filtering 

```{r}
mat[mat > 15]
```

This will return a vector of all the elements > 15 in the matrix `mat`


#### 7.1.6 Matrix summation [ELEMENTWISE]

```{r}
mat + mat
```

You can perform other operations like multiplication, subtraction, division.

#### 7.1.7 Matrix multiplication

```{r}
mat %*% mat
```
**This is not an element by element multiplication.**

### 7.2 Matrix operations [MATRIX FUNCTIONS]
#### 7.2.1 Sums
```{r}
colSums(prices.matrix)
rowSums(prices.matrix)
```

#### 7.2.2 Means
```{r}
rowMeans(prices.matrix)
colMeans(prices.matrix)
```

#### 7.2.3 `cbind()` and `rbind()` function to add columns and rows to an existing matrix
Let's use the fruit price example to see the application

```{r}
Grapes <- c(13, 14, 11, 10, 10)
fruit.prices <- rbind(prices.matrix, Grapes)
print(fruit.prices)
```

We can add a row that shows the weekly average price of the fruits

```{r}
avg <- rowMeans(fruit.prices)
prices.matrix <- cbind(fruit.prices, avg)
prices.matrix
```


# 8. Matrix Indexing
We start by creating a new matrix `mat`.
```{r}
mat <- matrix(1:20, byrow = T, nrow = 5)
mat
```

### 8.1 Index first row and all columns. Returns a vector

```{r}
mat[1,]
```

### 8.2 Just the Second column

```{r}
mat[,2]
```

### 8.3 First two rows and all columns

```{r}
mat[1:2,]
```

### 8.4 Last 2 rows and first 2 columns

```{r}
mat[4:5,1:2]
```

# 9. Data Frames
Data frames enable us to organize and mix data types to create powerful data structures,
unlike vectors and matrices where all the elements should be of the same data type.
Data frames can be thought of as an Excel spreadsheet.

### 9.1 Few build-in data frames in R

```{r}
state.x77
USPersonalExpenditure
women
```
To see all the built-in data sets use

```{r}
data()
```

### 9.2 Data frame functions
#### 9.2.1 Top 5 rows
```{r}
head(state.x77)
```

#### 9.2.2 Bottom 5 rows
```{r}
tail(state.x77)
```

#### 9.2.3 Structure of the data frame
```{r}
str(state.x77)
```

#### 9.2.4 Summary of Statistical Data
```{r}
summary(state.x77)
```

### 9.3 Create Data frames - Use the data.frame()
```{r}
days <- c('M','T','W','Th','F')
temp <- c(5,3,6,3,7)
rain <- c(T,F,F,T,F)

df <- data.frame(Days=days, Temperature=temp, Rain=rain)
df
```
**OR**
You can directly use the names of the vectors and they would be used as column names by default.
```{r}
df <- data.frame(days, temp, rain)
df
```
The structure and summary of the data frame `df` is. 
```{r}
str(df)
summary(df)
```
This is important during exploratory data analysis.  

### 9.4 Indexing and Sorting data frames
#### 9.4.1 Index the first row of `df`
```{r}
df[1,]
```

#### 9.4.2 Index the first column of `df`
```{r}
df[,1]
```

**OR**
```{r}
df$days         #This method returns a vector
```

**OR**
```{r}
df['days']      #This method returns a data frame
```


Instead of putting the index of the columns, we can use the names of the columns
```{r}
df[1:3, c('days','temp')]
```

#### 9.4.3 `subset()` function for doing a conditional subset  

Subset all rows where `rain == TRUE`  

```{r}
subset(df, subset = rain == T)
```

Subset all rows where `temp <= 22`
```{r}
subset(df, subset = temp <= 22)
```
The outputs are in data frame format.

#### 9.4.4 `order()` function
```{r}
sorted.temp <- order(df['temp'])
sorted.temp
```

This will output a vector of sorted temperature indexes. The default is ascending.
We can use these indexes to sort the data frame. 

```{r}
df[sorted.temp,]
```

To sort in descending order
```{r}
desc.temp <- order(-df['temp'])
df[desc.temp,]
```
**OR**
```{r}
desc.temp <- order(-df$temp)
df[desc.temp,]
```

To find the number of rows and number of columns use `nrow()` and `ncol()`
```{r}
nrow(df)
ncol(df)
```

To get the rownames and column names use `rownames()` and `colnames()` function
```{r}
colnames(df)
rownames(df)
```
### 9.5 Assigning cells a value
```{r}
df[[1,'temp']] <- 48
df
```
### 9.6 Referencing [Cells, Columns, Rows]
#### 9.6.1 Reference a cell in data frame using `[[]]`
```{r}
df[[1,2]]
```

**OR**
```{r}
df[[1,'temp']]
```


#### 9.6.2 Reference a row
```{r}
df[3,]
```

#### 9.6.3 Reference a column
Let's start with assigning mtcars dataset to a variable `df`
```{r}
df <- mtcars
```
To get the output in a vector format, there are 4 methods

**Method 1**
```{r}
df$mpg
```

**Method 2**
```{r}
df[,'mpg']
```

**Method 3**
```{r}
df[,1]
```

**Method 4**
```{r}
df[['mpg']]
```

To get the output in a data frame format
```{r}
df['mpg']
```

**OR**
```{r}
df[1]
```


Return multiple columns
```{r}
head(df[c('mpg', 'cyl')])
```

To add rows and columns to an existing data frame use the rbind() and cbind() functions
To add a new column you can use `df$colName <- 2*previous columns`

Rename columns using
`colnames(df) <- c("a","b","c","d","e","f")`

To rename a specific columns
`colnames(df)[1] <- "New_col"`

### 9.7 Filtering data
#### 9.7.1 Filtering rows as per condition
```{r}
mtcars[mtcars$mpg > 20, ]
```

#### 9.7.2 Filtering with multiple conditions
```{r}
mtcars[(mtcars$mpg > 20) & (mtcars$cyl == 6),]
```

#### 9.7.3 `subset()` function for filtering
```{r}
subset(mtcars, mpg > 20 & cyl == 6)
```
The subset function works well since you don't have to type `mtcars$name_of_column` everytime.  

### 9.8 Dealing with NA's
```{r}
is.na(mtcars)
```

The above function will return a data frame of logicals: TRUE if missing, FALSE O/W. 
```{r}
any(is.na(mtcars))
```

This will return a single logical value, False if no data is missing and True if 
atleast one data cell has missing values.


#### 9.8.1 To replace missing values in your entire data set use
```{r}
df[is.na(df)] <- 0
```

#### 9.8.2 To replace missing values in a specific column with the mean, you can do:
```{r}
mtcars$mpg[is.na(mtcars$mpg)] <- mean(mtcars$mpg)
```





